Predicting Transport in Lyotropic Liquid Crystal Membranes with Molecular Dynamics Simulations -- Outline

### 2017-08-29 Ben_Manuscripts/Paper1/Outline.tex
added results

  We chose initial guesses for the remaining structural parameters based on 
  experimental data and treated them as variables during model development.
  %BJC: I could potentially break this paragraph into smaller paragraphs centered around
  % the bolded structural parameters
  \begin{itemize}
    \item \textbf{The distance between pores} was based on experimental SAXS data for
    this system \cite{feng_thin_2016}.
    \item The model's pores are spaced 4.5 nm apart initially, $\approx$ 10 \% larger
    than the experimental value of 4.12 nm in order to reduce unintended repulsions 
    resulting from a tightly packed initial configuration.
    \item The layer spacing is based on experimental 2D WAXS data.
    \item The pattern shows reflections corresponding to features spaced 3.7 \angstrom apart.
    \item It has been hypothesized that the features are a result of pi-pi
    interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
    \item Our simulations tend to equilibrate to a wider interlayer spacing of
    $\approx$ 4.1 \angstrom, which inspired separate systems starting with layer 
    spacings greater than 4 \angstrom.
    \item We estimate the \textbf{pore radius} to be 0.6 nm based on past TEM images 
    and size exclusion rejection data\cite{feng_scalable_2014,feng_thin_2016,
    zhou_supported_2005}.
    \item Comparing a geometric measurement of pore size derived from an 
    atomistic model, to a less precise, experimentally derived pore size estimate,
    will give ambiguous results.
    \item What is meant by pore radius will not be clear until we have a clear picture
    of the nanoscopic pore environment.
    \item When constructing pores, we chose the carboxylate carbon from the monomer
    head group as a reference atom, and placed it a distance r from the pore center,
    where r is the pore radius.  %BJC: figure making that clear
    \item We will not make direct comparisons of pore radius between our model 
    and experiment to avoid the ambiguity. 
    \item The \textbf{relative interlayer orientation} was chosen based on clues from 
    diffraction data as well as the various known stacking modes of benzene 
    and substituted benzene rings: sandwiched, parallel-displaced and T-shaped
    ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
    \item The T-shaped configuration was ruled out based on the inconsistency of
    its $\approx$ 5 \angstrom equilibrium stacking distance ~\cite{sinnokrot_estimates_2002}.
    \item The system's preference towards the sandwiched vs. parallel displaced 
    stacking modes will be explored.
    %BJC: I think the following commentary belongs elsewhere, maybe even in the discussion
    \item Visualization of each configuration (\Cref{fig:sandwichedlayers,fig:offsetlayers})
    suggests entropic differences based on the way the tails are able to pack.
    \item In the sandwiched configuration, all tails start out directly on top
    of each other which may prevent closely stacked benzene rings.
    \item In the offset configuration, the tails are placed in between each other 
    which may allow layers to come together in a compact way.
    \item This difference may explain, in part, which stacking mode is more favorable.
  \end{itemize}

@@

>>>>

  We chose initial guesses for the remaining structural parameters based on 
  experimental data and treated them as variables during model development.
  %BJC: I could potentially break this paragraph into smaller paragraphs centered around
  % the bolded structural parameters
  \begin{itemize}
    \item \textbf{The distance between pores} was based on experimental SAXS data for
    this system \cite{feng_thin_2016}.
    \item The model's pore centers are spaced 4.5 nm apart initially, $\approx$ 10 \% larger
    than the experimental value of 4.12 nm in order to reduce unintended repulsions 
    resulting from a tightly packed initial configuration.
    \item To calculate the equilibrated pore spacing, we measured the distance between pore
    centers.
    \item Pore centers were located by averaging the coordinates of sodium ions in their 
    respective pores.
    \item Statistics were generated using the bootstrapping technique.
    \item For each bootstrap trial, we recreate an equilibrium trajectory by randomly 
    sampling from the original trajectory
    % This might be too much detail. Maybe leave excess detail for 
    \item We are interested in 5 pore-to-pore distances which should all be equal in 
    a perfect hexagonal array. Only 4 are independent  %BJC: small figure here for clarity? supplemental info?
    \item Each pore spacing has its own trajectory with its own average value
    \item The average value of each pore spacing is averaged to get the overall
    average pore spacing. 
    \item The standard deviation of average values is reported as the uncertainty in pore spacing.
    \item The \textbf{layer spacing} is based on experimental 2D WAXS data.
    \item The pattern shows reflections corresponding to features spaced 3.7 \angstrom apart.
    \item It has been hypothesized that the features are a result of pi-pi
    interactions between stacked aromatic rings \cite{feng_scalable_2014}. 
    \item Our simulations tend to equilibrate to a wider interlayer spacing of
    $\approx$ 4.1 \angstrom, which inspired separate systems starting with layer 
    spacings greater than 4 \angstrom.
    \item We estimate the \textbf{pore radius} to be 0.6 nm based on past TEM images 
    and size exclusion rejection data\cite{feng_scalable_2014,feng_thin_2016,
    zhou_supported_2005}.
    \item Comparing a geometric measurement of pore size derived from an 
    atomistic model, to a less precise, experimentally derived pore size estimate,
    will give ambiguous results.
    \item What is meant by pore radius will not be clear until we have a clear picture
    of the nanoscopic pore environment.
    \item When constructing pores, we chose the carboxylate carbon from the monomer
    head group as a reference atom, and placed it a distance r from the pore center,
    where r is the pore radius.  %BJC: figure making that clear
    \item We will not make direct comparisons of pore radius between our model 
    and experiment to avoid the ambiguity. 
    \item The \textbf{relative interlayer orientation} was chosen based on clues from 
    diffraction data as well as the various known stacking modes of benzene 
    and substituted benzene rings: sandwiched, parallel-displaced and T-shaped
    ~\cite{sinnokrot_estimates_2002} (\Cref{fig:sandwiched,fig:pd,fig:tshaped}).
    \item The T-shaped configuration was ruled out based on the inconsistency of
    its $\approx$ 5 \angstrom equilibrium stacking distance ~\cite{sinnokrot_estimates_2002}.
    \item The system's preference towards the sandwiched vs. parallel displaced 
    stacking modes will be explored.
    %BJC: I think the following commentary belongs elsewhere, maybe even in the discussion
    \item Visualization of each configuration (\Cref{fig:sandwichedlayers,fig:offsetlayers})
    suggests entropic differences based on the way the tails are able to pack.
    \item In the sandwiched configuration, all tails start out directly on top
    of each other which may prevent closely stacked benzene rings.
    \item In the offset configuration, the tails are placed in between each other 
    which may allow layers to come together in a compact way.
    \item This difference may explain, in part, which stacking mode is more favorable.
  \end{itemize}

@@

  \section*{Results and Discussion}

  \subsection*{Determination of Nanoscopic Structural Details}

  In order to construct an initial configuration which gives reliable 
  trends, we need to understand the composition of layers, how far apart
  to stack the layers, and how to orient them with respect to each other.
  \begin{itemize}
  	\item To verify our choices for each parameter, we compare our calculations
	to experimental small angle X-ray scattering (SAXS), wide angle X-ray
	scattering (WAXS), and ionic conductivity measurements.
  \end{itemize}

  To discern the composition of the monomer layers, we ran simulations of 
  systems created with 4 - 8 monomers per layer.
  \begin{itemize}
  	\item Both the sandwiched and parallel displaced configurations were tested.
	\item All systems remained stable for at least 50 ns.
	\item Table ~\ref{table:p2p} shows the pore spacing for all systems tested. 
	Systems built with 5 monomers in each layer equilibrate to a pore spacing
	that is most consistent with the experimental value of 4.12 nm derived from
	SAXS measurements (Figure~\ref{fig:SAXS}).
	\item We verified that the system stays partitioned into layers by plotting 
	pair correlation functions calculated between aromatic rings along the length
	of the pores (Fig~\ref{fig:zdf}).
	\item The remainder of this discussion will focus on the analysis of systems
	built with 5 monomers per layer. 
  \end{itemize}

  \begin{table}[H]
  \centering
  \begin{tabular}{ccc}
  \toprule
  		   & \multicolumn{2}{c}{Starting Configuration} \\
  \hline
  Monomers per layer & Sandwiched & Parallel Displaced \\
  \midrule
  4 & $3.71 \pm 0.04$ & $3.82 \pm 0.03$ \\
  5 & $4.20 \pm 0.04$ & $4.23 \pm 0.04$ \\
  6 & $4.83 \pm 0.03$ & $4.85 \pm 0.02$ \\
  7 & $4.73 \pm 0.03$ & $4.84 \pm 0.03$ \\
  8 & $5.08 \pm 0.04$ & $5.46 \pm 0.03$ \\
  \bottomrule
  \end{tabular}
  \caption{The pore spacing of the model increases as number of monomers
  in each layer increases. The pore spacing of a system starting in the 
  sandwiched configuration is systematically lower than that started in an 
  offset configuration. Systems built with 5 monomers per layer in a parallel
  displaced configuration result in a pore spacing closest to experimental
  data}\label{table:p2p} 
  \end{table}

  \begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{zdf5layered.png}
		\caption{}\label{fig:zdf_layered}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{zdf5offset.png}
		\caption{}\label{fig:zdf_offset}
	\end{subfigure}
	\caption{Pair distribution functions of aromatic carbons for the
	(a) 5 monomer per layer, sanwiched and (b) 5 monomer per layer, 
	parallel displaced configurations. Spikes in the distribution mark
	distinct layers. The magnitude of the spikes with respect to the 
	average suggest that the 5 monomer per layer, sandwiched configuration
	possesses the highest degree of layer partitioning.}\label{fig:zdf}
  \end{figure}

  The initial distance between layers can influence the approach towards
  equilibrium.

  Experimental WAXS measurements encode all structural details.

  \begin{figure}[ht]
	\centering
	\begin{subfigure}[t]{0.47\linewidth}
		\centering
		\raisebox{.2\textwidth}{%
		\includegraphics[width=\linewidth]{WAXS_soft_confined.png}
		}
		\caption{}\label{fig:WAXS}
	\end{subfigure}
	\begin{subfigure}[t]{0.43\linewidth}
		\centering
	%	\vspace{12mm}
		\includegraphics[width=\linewidth]{SAXS.png}
		\caption{}\label{fig:SAXS}
	\end{subfigure}
	\caption{(a) 2D wide angle X-ray scattering gives details about repeating
	features less than 1 nanometer apart. (b) 1D small angle X-ray scattering 
	indicates hexagonal packing of pores as well as the spacing between pores.}\label{fig:SWAXS}
  \end{figure}

  We varied the relative interlayer orientation between sandwiched and 
  parallel-displaced based on our knowledge of the stability of these two
  pi-stacking modes.

  \begin{figure}
	\centering
	\begin{subfigure}{0.475\textwidth}
		\hspace{-1.2cm}
		\centering
		\includegraphics[width=\textwidth,valign=t]{rzlayeredrestrained.png}
		\caption{}\label{fig:rzplayeredrestrained}
	\end{subfigure}
	\begin{subfigure}{0.475\textwidth}
		\hspace{-1.2cm}
		\centering
		\includegraphics[width=\textwidth,valign=t]{rzoffsetrestrained.png}
		\caption{}\label{fig:rzoffsetrestrained}
	\end{subfigure}
	\caption{(a) Simulated X-ray diffraction of a sandwiched configuration
	with restraints placed on aromatic carbons shows all major features
	present in experimental WAXS. Near solid lines at constant z are a result of 
	the highly ordered aromatic rings. (b) Simulated X-ray diffraction of a similarly
	restrained parallel displaced configuration may also contain all
	major experimental WAXS features. One can argue that R-spots is not present
	but it is difficult to distinguish because of its intersection with the solid
	R-helix line. Longer simulations are necessary to determine which structure
	is the best match to experiment}\label{fig:XRDrestrained}
  \end{figure}

  Full comparison of experimental 2D WAXS with simulated X-ray diffraction
  patterns produced from equilibrated MD trajectories shows the most consistency
  with the sandwiched configuration in Basin A.

  \begin{figure}[H]
	\centering
	\begin{subfigure}{0.31\textwidth}
		\centering
		\hspace{-0.9cm}
		\includegraphics[width=\textwidth]{sandwich_rzplot.png}
		\caption{}\label{fig:sandwich_rzplot}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\centering
		\hspace{-0.9cm}
		\includegraphics[width=\textwidth]{offset_rzplot.png}
		\caption{}\label{fig:offsetrzplot}
	\end{subfigure}
	\begin{subfigure}{0.31\textwidth}
		\centering
		\hspace{-0.9cm}
		\includegraphics[width=\textwidth]{disorder_rzplot.png}
		\caption{}\label{fig:disorder_rzplot.png}
	\end{subfigure}
	\caption{(a) All major features except R-helix are present in
	XRD patterns resulting from an equilibrated sandwiched configuration 
	in Basin A. R-pi intersescts R-alkanes. R-helix faded during equilbration.
	(b) All major features execpt R-spots are present in XRD patterns 
	resulting from an equilibrated parallel displaced configuration in Basin A.
	R-helix exists faintly in the expected location. (c) R-pores and R-alkanes
	are the only major features present in XRD patterns resulting from an 
	equilibrated disordered phase}\label{fig:XRDsim}
  \end{figure}

  \begin{wrapfigure}{R}{0.4\textwidth}
        \centering
        \includegraphics[width=0.4\textwidth]{tilt.png}
        \caption{The average angle between alkane chains and the xy plane is nearly zero degrees}\label{fig:tilt}
  \end{wrapfigure}

  The spots that appear in the simulated XRD pattern of the sandwiched
  conformation are a result of the way alkane tails pack together.

  \begin{figure}
	\centering
	\begin{subfigure}{0.45\textwidth}
		\centering
		\hspace{-1cm}
		\vspace{1cm}
		\includegraphics[width=\textwidth,scale=2]{tails_topview.png}
		\caption{}\label{fig:tails_topview}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{tails_rzplot.png}
		\caption{}\label{fig:tails_rzplot}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{centroids_box.png}
		\caption{}\label{fig:centroids}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\centering
		\includegraphics[width=\textwidth]{angles_traj_layered.png}
		\caption{}\label{fig:angle_distribution}
	\end{subfigure}
	\caption{(a) The trajectory can be stripped of all atoms except carbon
	atoms in monomer tails. (b) Simulated diffraction of the tail-only trajectory
	still gives rise to R-spots. (c) Finding the center of mass and visualizing 
	their coordinates reveals the hexagonal-like packing of the tails. (d) The 
	distribution created by measuring the angle between each centroid (e.g. red 
	in (c)) and its neareset neighbors (e.g. green in (c)) with respect to the xy
	plane has distinct spikes near 30\degree, which is consistent with the location
	of R-spots}\label{fig:tail_packing}
  \end{figure}

  Basin A and Basin B are two configurationally metastable basins.

  \begin{figure}
        \centering
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{280K_tramp_close.png}
                \caption{}\label{fig:280K_pore}
        \end{subfigure}
        \begin{subfigure}[b]{0.475\textwidth}
                \centering
                \includegraphics[width=\textwidth]{340K_tramp_close.png}
                \caption{}\label{fig:340K_pore}
        \end{subfigure}
        \vskip\baselineskip
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{p2p_tramp.png}
                \caption{}\label{fig:p2p_tramp}
        \end{subfigure}
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{thickness_tramp.png}

                \caption{}\label{fig:thickness_tramp}
        \end{subfigure}
        \begin{subfigure}[b]{0.325\textwidth}
                \centering
                \includegraphics[width=\textwidth]{order_tramp.png}
                \caption{}\label{fig:order_tramp}
        \end{subfigure}
        \caption{(a) The open pore structure exhibited by a structure equilibrated
	at 280K is characteristic of Basin A. (b) The closed pore structure with 
	a high degree of radial disorder exhibited when the structure in (a) is 
	heated to 340K is characteristic of Basin B. (c) A plot of distance between pores
	vs. temperature changes slope near 325K. (d) A plot of membrane thickness vs. 
	temperature changes slope near 325K. (e) The plot of the ratio of pore radius to
	its uncertainty changes slope near 315K.}\label{fig:tramp}
  \end{figure}

  \begin{figure}
	\centering
	\begin{subfigure}{0.325\textwidth}
		\centering
		\includegraphics[width=\textwidth]{p2p_layered.png}
		\caption{}\label{fig:p2p_layered}
	\end{subfigure}
	\begin{subfigure}{0.325\textwidth}
		\centering
		\includegraphics[width=\textwidth]{thickness_layered.png}
		\caption{}\label{fig:thickness_layered}
	\end{subfigure}
	\begin{subfigure}{0.325\textwidth}
		\centering
		\includegraphics[width=\textwidth]{order_layered.png}
		\caption{}\label{fig:order_layered}
	\end{subfigure}
	\caption{In all cases, blue lines represent the measured value of
	the order parameter, black lines are average values calculated from
	equilibrated Basin B systems at each temperature, green shaded regions
	represent the standard deviation of each of the black line values, and 
	red dashed lines show where the temperature is bumped to the next level.
	(a) The pore spacing decreases smoothly as it approaches Basin B values.
	(b) Membrane thickness increases steadily with temperature, however
	it is still far from the Basin B value after 200 ns at 320K. (c) The
	ratio of pore radius to is uncertainty rapidly decreases initially and
	approaches the Basin B value smoothly until they finally overlap at 320K.}\label{fig:phase_transition}
  \end{figure}

  \subsection*{Ionic conductivity calculation}

  The model gives reasonable estimates of ionic conductivity for both phases.

  \begin{table}[H]
  \centering
  \begin{tabular}{ccc}
  \toprule
  \multicolumn{3}{c}{Calculated Ionic Conductivity \si{\siemens\per\meter}} \\
  \hline
  Method & Basin A & Basin B \\
  \midrule
  Nernst Einstein & \num{1.23e-4} (0.01) & \num{1.76e-4} (0.02) \\
  Collective Diffusion & \num{1.40e-4} (0.32) & \num{4.6e-4}(2.4) \\
  Experiment & \num{1.33e-5} (0.10) & -- \\
  \bottomrule
  \end{tabular}
  \caption{Calculated ionic conductivity using Nernst-Einsten and Collective Diffusion 
  agree within error. Both methods give calculated values of ionic conductivity which
  are an order of magnitude higher than experimental values~\label{table:conductivity}}.
  \end{table}

  \subsection*{Implementation of the crosslinking algorithm}

  We applied the crosslinking algorithm to an equilibrated structure in Basin B.

  \begin{figure}[H]
	\centering
	\begin{subfigure}{0.31\textwidth}
		\centering
		\includegraphics[width=\textwidth]{p2p_diagram.PNG}
		\caption{}\label{fig:p2p_diagram}
	\end{subfigure}
		\begin{subfigure}{0.31\textwidth}
		\centering
		\includegraphics[width=\textwidth]{no_xlink_p2p.png}
		\caption{}\label{fig:no_xlink_p2p}
	\end{subfigure}
		\begin{subfigure}{0.31\textwidth}
		\centering
		\includegraphics[width=\textwidth]{xlink_p2p.png}
		\caption{}\label{fig:xlink_p2p}
	\end{subfigure}
	\caption{(a) The legends of the plots in (b) and (c) refer to the numbers shown.
	Each numbered circle represents a pore. Distances are measured along each of the 
	lines shown in addition to the distance from pore 1 to pore 4. (b) The positions
	of individual pores fluctuate in an uncrosslinked system. (c) The positions of
	individual pores in the crosslinked system are stable relative to the uncrosslinked
	system}\label{fig:xlink}
  \end{figure}
